---
title: "Confidence Intervals for the EXTEND Data"
subtitle: "Analysis using main effects and noise variables"
author: "Jeremiah Jones"
date: "2022-10-3"
output:
  pdf_document: 
    keep_tex: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
pretty_list_english <- function(ch_vec) {
  n <- length(ch_vec)
  ret <- if(n <= 0) {
    stop("Degenerate length!")
  } else if(n == 1) {
    ch_vec
  } else if(n == 2) {
    paste0(ch_vec[1], " and ", ch_vec[2])
  } else {
    last <- ch_vec[n]
    first <- paste(ch_vec[-n], collapse=", ")
    paste0(first, ", and ", last)
  }
  ret
}
collapse_noise <- function(ch_vec) {
  if(length(ch_vec) <= 1) {
    stop("Only length 1!")
  }
  n <- length(ch_vec)
  noisesB <- stringr::str_detect(ch_vec, "noise")
  out_vec <- ch_vec[!noisesB]
  c(out_vec, paste0(sum(noisesB), " noise variables"))
}
list_the_model_vars <- function(stage) {
  ret <- if(stage == 2) {
    colnames(Q2_sel_mat) %>% stringr::str_remove("a2c:") %>% 
      collapse_noise %>% pretty_list_english
  } else if(stage == 1) {
    colnames(Q1_sel_mat) %>% stringr::str_remove("a1c:") %>% 
      collapse_noise %>% pretty_list_english
  } else {
    stop("Incorrect Stage!")
  }
  ret
}
Pval_format <- function(p){
  ret <- if(p >= .01){
    sprintf("%.2f", p)
  } else {
    "<.01"
  }
  ret
}
```

```{r load, include=FALSE}
load("../results/run.RData")
```


We analyzed the Extend data using our methods. We chose the first five variables selected by LAR as the model selection mechanism in both stages to compare our proposed UPOSI-based method with selective inference. In addition to these two inference techniques, we also included naive methods that ignored the model selection event: one based on the robust sandwich estimator, and another based on the perturbation bootstrap with standard exponential weights.

Since the data is relatively low-dimensional, we simulated 50 additional baseline variables as i.i.d. standard Normal variates. These have no relationship to the outcome or any additional variables, yet may be spuriously selected by the model selection mechanism, which specifies a model size *a priori.* Consequently, we use these noise variables to judge the anticonservativeness of each of the four inference methods. The "full model" in each stage includes only the "main effects" of the variables observed prior to that stage's decision, including the 50 simulated based variables. In total, that leads to a full model of size `r ncol(Q2_big)` in the second stage and `r ncol(Q1_mat)` in the first.

### Stage 2

Using the previously-described LAR selection method, the Stage 2 model included `r list_the_model_vars(2)` as the tailoring variables. The inferential results are summarized below:


```{r stage2, message=FALSE, warning=FALSE, echo=FALSE}
library(dplyr)
library(kableExtra)
boot2 <- tibble(
  method = "Naive bootstrap",
  var = rownames(ci_stg2_boot),
  coefficient=coef(rob_Q2_fit),
  lower=ci_stg2_boot[,1],
  upper=ci_stg2_boot[,2],
  length=ci_stg2_boot[,3]
)
sand2 <- tibble(
  method = "Naive Sandwich Estimator",
  var = rownames(ci_sand_stg2),
  coefficient=coef(rob_Q2_fit),
  lower=ci_sand_stg2[,1],
  upper=ci_sand_stg2[,2],
  length=ci_sand_stg2[,3]
)
si2 <- tibble(
  method = "SI",
  var = rownames(ci_si_stg2),
  coefficient=coef(rob_Q2_fit),
  lower=ci_si_stg2[,1],
  upper=ci_si_stg2[,2],
  length=ci_si_stg2[,3]
)
uposi2 <- tibble(
  method = "UPOSI",
  var = rownames(ci_stg2),
  coefficient=coef(rob_Q2_fit),
  lower=ci_stg2[,1],
  upper=ci_stg2[,2],
  length=ci_stg2[,3]
)


boot2 %>% union(si2) %>% union(uposi2) %>% union(sand2) %>%
  select(var, method, everything()) %>% 
  mutate(var = stringr::str_remove(var, "a2c:")) %>% 
  arrange(var, method) %>% 
  rename(Variable=var,
         `Inference Method`=method,
         Coefficient=coefficient,
         `Lower Limit`=lower, `Upper Limit`=upper,
         Length=length
         ) %>% 
  kbl(digits=3) %>%
  # kbl(digits=3, booktabs = T) %>% 
  kable_paper("striped", full_width=F) %>%
  add_header_above(c(" "=3, "Confidence Intervals"=3)) %>% 
  collapse_rows(columns=1, valign="top")

```

```{r overall-test-2, echo=FALSE}
summ2 <- summary(rob_Q2_fit)
Pval_par2 <- 1-pf(summ2$fstatistic[1], summ2$fstatistic[2], summ2$fstatistic[3])
```

Compared to the SI intervals, our proposed intervals tend to be smaller in
length.
In fact, one of the SI intervals has infinite length while our proposed interval is finite.
Inference that is unadjusted for the selection process concludes that three of the selected variables are significantly different from zero, including one of the two selected noise variables.

The overall F-test provided by the `summary.lm` function leads to an overall P-value of `r Pval_format(Pval_par2)`, concluding that at least one of the selected coefficients is significantly different from zero at the .05 level. On the other hand, the omnibus test based on the UPoSI statistic leads to a P-value of `r Pval_format(Pval2)`, leading the researcher to avoid making any conclusions regarding the measured covariates.


```{r results='asis', include=FALSE}
boot2 %>% union(si2) %>% union(uposi2) %>% union(sand2) %>%
  select(var, method, everything()) %>% 
  mutate(var = stringr::str_remove(var, "a2c:")) %>% 
  arrange(var, method) %>% 
  rename(Variable=var,
         `Inference Method`=method,
         Coefficient=coefficient,
         `Lower Limit`=lower, `Upper Limit`=upper,
         Length=length
         ) %>% 
  # kbl(digits=3) %>%
  kbl(digits=3, format="latex", booktabs = T) %>%
  # kable_paper("striped", full_width=F) %>%
  add_header_above(c(" "=3, "Confidence Intervals"=3)) %>% 
  collapse_rows(columns=1, valign="top")
```

```{r include=FALSE}
summary(rob_Q2_fit) -> foo
1-pf(foo$fstatistic[1], foo$fstatistic[2], foo$fstatistic[3])
print(foo)
Pval2
```


### Stage 1

Applying these methods to the Stage 1 problem, the selected model included `r list_the_model_vars(1)` as the tailoring variables. The inferential results are summarized below:

```{r stage1, message=FALSE, warning=FALSE, echo=FALSE}
boot1 <- tibble(
  method = "Naive bootstrap",
  var = rownames(ci_stg1_boot),
  coefficient=coef(rob_Q1_fit),
  lower=ci_stg1_boot[,1],
  upper=ci_stg1_boot[,2],
  length=ci_stg1_boot[,3]
)
sand1 <- tibble(
  method = "Naive Sandwich Estimator",
  var = rownames(ci_sand_stg1),
  coefficient=coef(rob_Q1_fit),
  lower=ci_sand_stg1[,1],
  upper=ci_sand_stg1[,2],
  length=ci_sand_stg1[,3]
)
si1 <- tibble(
  method = "SI",
  var = rownames(ci_si_stg1),
  coefficient=coef(rob_Q1_fit),
  lower=ci_si_stg1[,1],
  upper=ci_si_stg1[,2],
  length=ci_si_stg1[,3]
)
uposi1 <- tibble(
  method = "UPOSI",
  var = rownames(ci_stg1),
  coefficient=coef(rob_Q1_fit),
  lower=ci_stg1[,1],
  upper=ci_stg1[,2],
  length=ci_stg1[,3]
)


boot1 %>% 
  union(si1) %>%
  union(uposi1) %>% union(sand1) %>%
  select(var, method, everything()) %>% 
  mutate(var = stringr::str_remove(var, "a1c:")) %>% 
  arrange(var, method) %>% 
  rename(Variable=var,
         `Inference Method`=method,
         Coefficient=coefficient,
         `Lower Limit`=lower, `Upper Limit`=upper,
         Length=length
         ) %>% 
  kbl(digits=3) %>%
  # kbl(digits=3, booktabs = T) %>%
  kable_paper("striped", full_width=F) %>%
  add_header_above(c(" "=3, "Confidence Intervals"=3)) %>% 
  collapse_rows(columns=1, valign="top")

```


```{r overall-test-1, echo=FALSE, results='hide'}
summ1 <- summary(rob_Q1_fit)
Pval_par1 <- 1-pf(summ1$fstatistic[1], summ1$fstatistic[2], summ1$fstatistic[3])
```

The naive inference methods find significant relationships for 3 out of the 4 selected synthetic variables, demonstrating the anticonservativeness properties of this inference technique. As in the second stage, the proposed method leads to smaller confidence intervals when compared to SI, avoids the infinite-length issues of the latter, and avoids the anticonservativeness of the naive methods.

Using the parametric F-test of overall significance after ignoring the selection process, we arrive at a highly-significant P-value of `r Pval_format(Pval_par1)`. On the other hand, the proposed overall test has a P-value of `r Pval_format(Pval1)`, leading researchers to avoid making claims of significance derived from performing selection on this data.

```{r}
boot2 %>% union(si2) %>% union(uposi2) %>% #union(sand2) %>%
  mutate(stage=2) %>% 
  mutate(var = stringr::str_remove(var, "a2c:")) %>% 
  union(
    boot1 %>% 
      union(si1) %>%
      union(uposi1) %>% #union(sand1) %>%
      mutate(stage=1, 
             var = stringr::str_remove(var, "a1c:"))
  ) %>% 
  select(stage, var, coefficient, method, everything()) %>% 
  arrange(stage, var, method) %>% 
  mutate(method=if_else(method=="Naive bootstrap", "Naive", method),
         method = if_else(sign(lower)*sign(upper)>0, paste0(method, "*"), method)) %>% 
  rename(Variable=var,
         `Inference Method`=method,
         Coefficient=coefficient,
         `Lower Limit`=lower, `Upper Limit`=upper,
         Length=length,
         Stage=stage
         ) %>% 
  # kbl(digits=3) %>%
  kbl(digits=3, booktabs = T) %>%
  # kable_paper("striped", full_width=F) %>%
  add_header_above(c(" "=4, "Confidence Intervals"=3)) %>% 
  collapse_rows(columns=1:3, valign="middle")

```

```{r}
library(ggplot2)
plot_df <- boot2 %>% union(si2) %>% union(uposi2) %>% #union(sand2) %>%
  mutate(stage=2) %>% 
  mutate(var = stringr::str_remove(var, "a2c:")) %>% 
  union(
    boot1 %>% 
      union(si1) %>%
      union(uposi1) %>% #union(sand1) %>%
      mutate(stage=1, 
             var = stringr::str_remove(var, "a1c:"))
  ) %>% 
  select(stage, var, coefficient, method, everything()) %>% 
  arrange(stage, var, method) %>% 
  mutate(method=if_else(method=="Naive bootstrap", "Naive", method),
         sig = factor(sign(lower)*sign(upper)<=0, labels=c("Discovery", "No Discovery"), levels=c(F,T))) %>% 
  mutate(stage=factor(stage, levels=c(2,1), labels=c("Stage 2", "Stage 1")))
```

```{r discovery-rate}
plot_df %>% 
  ggplot(aes(x=method, y=var)) + 
  # geom_point(aes(color=sig, shape=sig, fill=sig), size=3) +
  geom_tile(aes(fill=sig), color="black") +
  scale_fill_brewer(type="div", palette=3) +
  scale_x_discrete(expand=c(0,0)) +
  scale_y_discrete(expand=c(0,0)) +
  facet_grid(rows=vars(stage), scales = "free_y") +
  theme_bw() + 
  theme(legend.position = "bottom",
        # panel.grid = element_blank(),
        # panel.border = element_blank()
        ) +
  labs(x="", y="", title="Discoveries by Inference Method",
       fill="", color="", shape="")
```

```{r confint-length}
plot_df %>% 
  ggplot(aes(y=var, x=length)) + 
  geom_point(aes(color=method, shape=method), size=3) +
  scale_x_continuous(trans="log2", oob=scales::censor,
                     # limits=4**c(-2, 1.6),
                     # breaks=4**c(-2:2),
                     # labels = sapply(c(-2:2), labeling)
                     labels = scales::trans_format("log2", scales::math_format(2^.x))
                     ) +
  # scale_fill_brewer(type="qual", palette=1) +
  scale_color_brewer(type="qual", palette=1) +
  # scale_shape_manual(values=c(18,17,20)) +
  # scale_shape_manual(values = c(1,2)) +
  facet_grid(rows=vars(stage), scales = "free_y") +
  theme_bw() + theme(legend.position = "bottom") +
  labs(x="Confidence Interval Length", y="Term", title="", color="Inference Method", shape="Inference Method")
```

